---
title: "Web APIs"
author: "Jered Ataky"
date: "2020-10-23"
output: 
  openintro::lab_report: default
  html_document:
    number_sections: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

<style>
div.aquamarine { background-color:#7fffd4; border-radius: 10px; padding: 5px;}
</style>
<div class = "aquamarine">

**Problem**

The New York Times web site provides a rich set of APIs, as described here: https://developer.nytimes.com/apis
I need to start by signing up for an API key.
My task is to choose one of the New York Times APIs, construct an interface in R to read in the JSON data, and
transform it into an R DataFrame.

**Approach**

The New York Times API that interested me the most was "Books API".
After requesting the API key, I was able to access it in R. 
I extracted the data of "hardcover fiction books" from the books API, and this was in JSON format. 
I then used "fromjson" to convert the JSON data to R objects.

Since "fromjson" converts the JSON data and includes the list of the object related to it,
I retrieved only the data frame containing all the books info that I need.

</div> \hfill\break

## Packages

```{r load-packages, message=FALSE}

library(httr)
library(jsonlite)
library(tidyverse)
library(kableExtra)

```


## Read the API

```{r}
# Read JSON data from the API with the key

books <-"https://api.nytimes.com/svc/books/v3/lists/current/hardcover-fiction.json?api-key=s6WdQYAxrWRmhzTGoytPJplwgWujefbc"

```

## Convert R objects from JSON

```{r}

# Transform JSON data into an R DataFrame
# Retrieve the data frame of books data

df <- fromJSON(books)[[5]][[11]]
dim(df)
```

```{r}

# Get the insights

glimpse(df)
```



```{r}

# New DataFrame  with some necessary columns from the original df
# (for analysis purpose)

df1 <- df[c("rank", "publisher", "title", "author", "primary_isbn13")]

df1 %>%
  kbl(caption = "Hardcover fiction books") %>%
  kable_material(c("striped", "hover")) %>%
  row_spec(0, color = "indigo")

```



## Extra

```{r}

# Publisher df. Order books by publisher

df_pub <- df1 %>%
  group_by(publisher) %>%
  summarise(books_published = n())

df_pub <- df_pub[order(-df_pub$books_published), ]


# Plot the ranked books by publisher
# (Visualize which publisher has more books in the ranking)

df_pub %>%
  
  ggplot(aes(reorder(publisher, books_published), books_published)) +
  
  geom_col(aes(fill = books_published)) +
  
  scale_fill_gradient2(low = "yellow",
                       high = "purple",
                       midpoint = median(df_pub$books_published)) +

  
coord_polar() +

  
  labs(title = "Ranked hardcover fiction books by publisher", x = NULL, y = NULL)  
  

```

## Findings


<style>
div.aquamarine { background-color:#7fffd4; border-radius: 10px; padding: 5px;}
</style>
<div class = "aquamarine">



If you have access to it, API makes it easier on getting directly the data you'd want...

</div> \hfill\break




