---
title: "Problem 2: Happiness Vs GDP"
author: "Jered Ataky"
date: "`r Sys.Date()`"
output: 
  openintro::lab_report: default
  html_document:
    number_sections: yes
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Background


<style>
div.aquamarine { background-color:#7fffd4; border-radius: 10px; padding: 5px;}
</style>
<div class = "aquamarine">

This dataset was posted by Magnus Skonberg on week 5 discussion board in DATA 607.
The entire document about it can be found in this link:
https://www.kaggle.com/unsdsn/world-happiness

The proposed analysis is: 

Compare the (Happiness) Score and GDP per capita for the Top 20 countries 
to observe the correlation (if there is any).

Thus I will based my analysis in 2019 data.

</div> \hfill\break


## Libraries

```{r}

library(tidyverse)
library(kableExtra)
library(Hmisc)
```


## Data Transformation


### Data insights and cleansing 

```{r}
# Get the data

data <- read.csv("https://raw.githubusercontent.com/jnataky/DATA-607/master/A2_Various_dataset_transformation/happy_data2019.csv")

```

```{r}
# make a copy of the data frame 

data_copy <- data

dim(data_copy)

```

```{r}

# Get the insights

head(data_copy)

```

```{r}
# Check for missing values

sum(is.na(data_copy))

```


```{r}
# Retrieve columns names

colnames(data_copy)

```

```{r}
# Rename columns

data_copy <- data_copy %>%
  rename(rank = Overall.rank, country = Country.or.region, happiness_score = Score, GDP_per_capita =  GDP.per.capita, freedom = Freedom.to.make.life.choices, generosity = Generosity, corruption = Perceptions.of.corruption)

```

### Tidying data 

```{r}
# make a copy of data set
data_copy1 <- data_copy

```


```{r}
# create new data frame with necessary variables for analysis

data1 <- data_copy1 %>%
  select(country, happiness_score, GDP_per_capita)

```

```{r}
# select top 20 countries

data2 <- head(data1, 20)
```

```{r}
data2 %>%
  kbl(caption = "Countries rank for Happiness & GDP per capita", align = 'c') %>%
  kable_material(c("striped", "hover")) %>%
  row_spec(0, color = "indigo")

```


## Data Analysis

```{r}

ggplot(data = data2) +
  geom_point(mapping = aes(x = GDP_per_capita, y = happiness_score, color = country)) +
  geom_smooth(mapping = aes( x = GDP_per_capita, y = happiness_score), se = FALSE, color = "red")


```

There is not a linear relationship between GDP_per_capita and happiness
or it is just a very small association between GDP per capita and
happiness.

Let calculate to verify this by calculation:

```{r}

cor(data2$GDP_per_capita, data2$happiness_score)

```


## Findings


<style>
div.aquamarine { background-color:#7fffd4; border-radius: 10px; padding: 5px;}
</style>
<div class = "aquamarine">


As Rousseau write in 1750: "Money buys everything, except morality and citizens."
Used ironically as "MONEY CAN'T BUY HAPPINESS"...   

We can confirm from this analysis that money can't buy happiness as there is very 
small association between GDP per capita and happiness.



</div> \hfill\break




