---
title: "Problem 1: Student Performance"
author: "Jered Ataky"
date: "`r Sys.Date()`"
output: 
  openintro::lab_report: default
  html_document:
    number_sections: yes
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Background


<style>
div.aquamarine { background-color:#7fffd4; border-radius: 10px; padding: 5px;}
</style>
<div class = "aquamarine">

This dataset was posted by Jered Ataky on week 5 discussion board in DATA 607.
The entire document about it can be found in this link:
https://www.kaggle.com/spscientist/students-performance-in-exams

The proposed analyses were:

- Correlation between students performance and parental level of education.

-if scores can be predicted based on the other variables such as test preparation, 
parental level of education, and lunch cost. We can see if these variables affect the scores and build a model.  

-We can also see if some scores are correlated with each other, such as reading and writing. 


</div> \hfill\break


## Libraries

```{r}

library(tidyverse)
library(kableExtra)
```

## Data Transformation

### Data insights and cleansing 

```{r}
# Get the data

data <- read.csv("https://raw.githubusercontent.com/jnataky/DATA-607/master/A2_Various_dataset_transformation/students_performance.csv")

```

```{r}
# make a copy of the data frame 

data_copy <- data

dim(data_copy)

```

```{r}
head(data_copy)

```


```{r}
# Retrieve columns names

colnames(data_copy)

```


```{r}
# Check for missing values

sum(is.na(data_copy))

```
```{r}
# Rename columns

data_copy <- data_copy %>%
  rename(parent_level = parental.level.of.education, lunch = lunch, prep_test = test.preparation.course, math =  math.score, reading = reading.score, writing = writing.score)

head(data_copy)


```

### Tidying the data frame

#### Subset data frame in 3 for further analysis

```{r}
# another copy

data_copy1 <- data_copy

```


```{r}

# Subset for parent_level and tests score

data_1 <- data_copy1[c(3,6:8)]
head(data_1)
```

```{r}
# Check different parent_level

unique(data_1$parent_level)

```

```{r}
# Gather columns

data_1 <- data_1 %>%
  gather("test", "score", 2:4)

```

```{r}
# Wider the data frame

data_1 <- data_1 %>%
  group_by(parent_level, test) %>%
  summarise(average_score = round(mean(score), 0))


```

```{r}

# Kable for tidy table

data_1 %>%
  kbl(caption = "Test score mean with parent level of education", align = 'c') %>%
  kable_material(c("striped", "hover")) %>%
  row_spec(0, color = "indigo")

```


```{r}

# Subset for lunch and tests score

data_2 <- data_copy1[c(4,6:8)]
```

```{r}
# Check different type of lunch

unique(data_2$lunch)

```

```{r}
# Gather columns

data_2 <- data_2 %>%
  gather("test", "score", 2:4)

```

```{r}
# Wider the data frame

data_2 <- data_2 %>%
  group_by(lunch, test) %>%
  summarise(average_score = round(mean(score), 0))

```

```{r}

# Kable for tidy table

data_2 %>%
  kbl(caption = "Test score mean with type of lunch offered to students", align = 'c') %>%
  kable_material(c("striped", "hover")) %>%
  row_spec(0, color = "indigo")

```


```{r}

# Subset for test_prep and tests score

data_3 <- data_copy1[c(5:8)]
```


```{r}
# Gather columns

data_3 <- data_3 %>%
  gather("test", "score", 2:4)

```


```{r}
# Wider the data frame

data_3 <- data_3 %>%
  group_by(prep_test, test) %>%
  summarise(average_score = round(mean(score), 0))

```


```{r}

# Kable for tidy table

data_3 %>%
  kbl(caption = "Test score mean with type of lunch offered to students", align = 'c') %>%
  kable_material(c("striped", "hover")) %>%
  row_spec(0, color = "indigo")

```


## Data Analysis

### Correlation between parental level of education and tests score

```{r}

ggplot(data = data_1) +
  geom_bar( mapping = aes(x = reorder(parent_level, average_score), y = average_score, fill = test), position = "dodge", stat = "identity") +
  facet_wrap(~ test, nrow = 3)

```

There is a positive correlation between parent level of education 
and students performance.


### Correlation between lunch and tests score


```{r}

ggplot(data = data_2) +
  geom_bar( mapping = aes(x = reorder(lunch, average_score), y = average_score, fill = test), position = "dodge", stat = "identity", width = 0.5) +
  facet_wrap(~ test, nrow = 3)

```
Students with standard lunch perform better than students with free/reduced lunch 


### Correlation between test preparation and tests score

```{r}

ggplot(data = data_3) +
  geom_bar( mapping = aes(x = reorder(prep_test, average_score), y = average_score, fill = test), position = "dodge", stat = "identity", width = 0.5) +
  facet_wrap(~ test, nrow = 3)

```
There is a positive correlation between completing test preparation and not completing
for all the 3 tests.


## Findings


<style>
div.aquamarine { background-color:#7fffd4; border-radius: 10px; padding: 5px;}
</style>
<div class = "aquamarine">


As we can see from analyses above, students performance are correlated to each of these three 
factors: parent level of education, lunch, and test preparation.

For parent level of education, the more the parents have a higher education 
the more students performed on all of the tests. This is maybe certain students tend to 
challenge themselves to their parents. Also something interesting is that 
the average performance for students whose parents level is "some high school" perform
better than students whose parents completed high school.

In regards to lunch, seems like parent income affect students performance in tests.

For test preparation, this is sort of expected that students preparing for tests 
perform better although we can find some particular cases out there.



</div> \hfill\break
