---
title: "Problem 3: World food-feed production"
author: "Jered Ataky"
date: "`r Sys.Date()`"
output: 
  openintro::lab_report: default
  html_document:
    number_sections: yes
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Background


<style>
div.aquamarine { background-color:#7fffd4; border-radius: 10px; padding: 5px;}
</style>
<div class = "aquamarine">

This dataset was posted by Rachel Greenlee on week 5 discussion board in DATA 607.
The data set includes food and feed production by country and food item 
from 1961 to 2013, including geocoding. 
The entire document about it can be found in this link:
https://www.kaggle.com/dorbicycle/world-foodfeed-production

The proposed analyses (taken and modified from the post): 


-Focus in on a specific country and see change over time over items .

-Compare food and feed on an item.


</div> \hfill\break


## Libraries

```{r}

library(tidyverse)
library(kableExtra)
```


## Data Transformation


### Data insights and cleansing 

```{r}
# Get the data

raw_data <- read.csv("https://raw.githubusercontent.com/jnataky/DATA-607/master/A2_Various_dataset_transformation/FAO.csv")

```

```{r}
# make a copy of the data frame 

copy_1 <- raw_data

# Get the insights

head(copy_1)

```



```{r}
# Retrieve columns names

colnames(copy_1)

```

```{r}
# New data set for only china

china_data <- copy_1 %>%
  filter(Area.Abbreviation == "CHN")

```

```{r}

# Consider top 3 products

china_data1 <- china_data %>%
  head(3)

```

```{r}

# Drop unnecessary columns

china_data2 <- china_data1 [-c(2:4, 6, 8:10)]

```


### Tidying the data frame

```{r}

# Make a copy of data frame

china1 <- china_data2


# Get rid of Area abbreviation

china1 <- china1[-c(1)]
# Gather the data

china1 <- china1 %>%
  gather("Year", "n_production", 3:55)

  
```

#### Subset data frame in 3 for further analysis

##### Wheat and products

```{r}

china1_w <- china1 %>%
  filter(Item == "Wheat and products")

# Arrange Year variable

china1_w$Year <- seq(1961,2013)

```


```{r}

# Kable for tidy table

china1_w %>%
  kbl(caption = "China' Number of production of Wheat and products from 1961 to 2013", align = 'c') %>%
  kable_material(c("striped", "hover")) %>%
  row_spec(0, color = "indigo")

```


#### Rice (Milled Equivalent)

```{r}

china1_r <- china1 %>%
  filter(Item == "Rice (Milled Equivalent)")

#Gather data

china1_r <- china1_r %>%
  pivot_wider(names_from = Element, values_from = n_production)


# Arrange Year variable

china1_r$Year <- seq(1961,2013)

```

```{r}

# Kable for tidy table

china1_r %>%
  kbl(caption = " Tab2. China' Number of production of rice (Milled equivalent) from 1961 to 2013", align = 'c') %>%
  kable_material(c("striped", "hover")) %>%
  row_spec(0, color = "indigo")

```


## Data Analysis

###  Wheat and products food production over time 

```{r}

# Plot Wheat and products food production over time 

ggplot( data = china1_w) +
  geom_line( mapping = aes(x = Year, y = n_production), color = "red") +
  labs(title ="China' Wheat and products food production over time")


```

Wheat and products have been increasing signifantly over the time


###  Rice (Milled equivalent) food production over time 

```{r}

# Plot Wheat and products food production over time 

ggplot( data = china1_r) +
  geom_line( mapping = aes(x = Year, y = Food), color = "red") +
  labs(title ="China' rice (Milled equivalent) production over time")


```
Rice (food) production has been wavering and increasing over time. 
(Rice feed not even produced (See Tab2) over the time). 


## Take Away


<style>
div.aquamarine { background-color:#7fffd4; border-radius: 10px; padding: 5px;}
</style>
<div class = "aquamarine">

In China, the wheat and products production over time has increased and
seems to develop more in years to come. In the other side, rice who was a big 
production for China around late 60s to around early 90s started decreasing 
tremendously in the point where it reached the lower production in the past
50 years. While rice (food), still have some production, the feed doesn't have any
production since 1979 (see Tab2).



</div> \hfill\break