---
title: "DATA607 ASS2: SQL and R"
author: "Jered Ataky"
date: "9/11/2020"
output: 
  openintro::lab_report: default
  html_document:
    number_sections: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Introduction


<style>
div.aquamarine { background-color:#7fffd4; border-radius: 10px; padding: 5px;}
</style>
<div class = "aquamarine">


I first created a survey of 6 movies which I have 5 persons responding.
After saving the survey in csv, I stored the results in MySQL database.


</div> \hfill\break



## Approach


<style>
div.aquamarine { background-color:#7fffd4; border-radius: 10px; padding: 5px;}
</style>
<div class = "aquamarine">


I will analyze the rating of each movie to know what is the most rated movie.
Then, go through each friend's rating of individual movies and overall rating 
of movies. This will help me to know what type of the movie the friend likes 
and if he or she is a fun of movies. Then, after I will recommend a movie 
if necessary. 

</div> \hfill\break


```{r}

# Load the results from MySQL to R
library(tidyverse)
library(RMySQL)

```

```{r MSQL}

# Accessing MySQL data base with password stored in a settings file

#rmariadb.settingsfile <- "C:\\data\\credentials.cnf"
#rmariadb.db <- "credentials"
#mydb <- dbConnect(RMariaDB::MariaDB(),default.file = rmariadb.settingsfile,group = rmariadb.db)

mydb <-dbConnect(RMariaDB::MariaDB(), user='root', password='mypassword', dbname ='labs', host='localhost')

```

```{r}
# Store the result in a dataframe movies

movies<- dbGetQuery(mydb, "SELECT * FROM movies")
```

```{r populate-data-frame}
library(tidyverse)
glimpse(movies)

```

## Handle missing values

```{r missing-data}
# Replace all missing values (a move not seen by a person) by 0

movies[is.na(movies)] = 0
```


```{r summarise-movies}
summary(movies)

```
## Best Movie by Average rating


```{r movies-rating-average}

movies %>%
  group_by(Movie_title) %>%
  summarise(mean_rating = mean(Rating)) %>%
  arrange(desc(mean_rating))
  
```


```{r movies-rating-average-plot}

movies %>%
  group_by(Movie_title) %>%
  summarise(mean_rating = mean(Rating)) %>%
  ggplot(., aes(y = Movie_title, x= mean_rating, fill = mean_rating)) + geom_bar(stat = 'identity')
  
```


## Individual rating by movie

```{r Batila-preference}

# Batila's preference

batila_movie <- movies %>%
  filter(Friend_name == "Batila")

ggplot(data = batila_movie, aes( y = Movie_title, x = Rating, fill = Rating)) + geom_bar(stat = 'identity')



```

```{r Musengie-preference}

# Musengie's preference

musengie_movie <- movies %>%
  filter(Friend_name == "Musengie")

ggplot(data = musengie_movie, aes( y = Movie_title, x = Rating, fill = Rating)) + geom_bar(stat = 'identity')



```



```{r Jonathan-preference}

# Jonathan's preference

jonathan_movie <- movies %>%
  filter(Friend_name == "Jonathan")

ggplot(data = jonathan_movie, aes( y = Movie_title, x = Rating, fill = Rating)) + geom_bar(stat = 'identity')



```



```{r Herve-preference}

# Herve's preference

herve_movie <- movies %>%
  filter(Friend_name == "Herve")

ggplot(data = herve_movie, aes( y = Movie_title, x = Rating, fill = Rating)) + geom_bar(stat = 'identity')


```


## MaV's preference

```{r MaV-preference}
MaV_movie <- movies %>%
  filter(Friend_name == "MaV")

ggplot(data = MaV_movie, aes( y = Movie_title, x = Rating, fill = Rating)) + geom_bar(stat = 'identity')


```
## Each friend's overall rating

```{r}

movies %>%
  group_by(Friend_name) %>%
  summarise(mean_rating = mean(Rating)) %>%
  ggplot(., aes(y = Friend_name, x= mean_rating, fill = mean_rating)) + geom_bar(stat = 'identity')
```



## Conclusion & recommendation

<style>
div.aquamarine { background-color:#7fffd4; border-radius: 10px; padding: 5px;}
</style>
<div class = "aquamarine">


I realized that Batila has watched all the  ovies and his overall rating 
for those movies were very high. Nothing I can recommend to him from my 
list of six movies. Musengie had watched 4 out 6 movies but her rating is kind
of the average so I would to recommend a very good movie to her.
When it comes to movie, "American son" had an average low rating just because 
it was watched only by two two people although the friends who'd seen, like it.
It is a good option to recommend to others.

</div> \hfill\break



```{r}
library(kableExtra)

# Construct a table frame for friends and recommended movies

f <- c("Musengie", "Jonathan", "Herve", "MaV")
m <- c("Six Underground", "American son", "American son", "Spencer confidential")


df <- data.frame(f,m)


colnames(df) <- c("Friend's name", "Recommended Movie")


  
# Movie recomendation table

df %>%
  kbl(caption = "Movie recommendation", align = 'c') %>%
  kable_material(c("striped", "hover")) %>%
  row_spec(0, color = "indigo")

```




