---
title: "DATA607 - Assignment 3"
author: "Jered Ataky"
date: "9/12/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

This assignment is about R character Manipulation and Date Processing.
These problems are constructed to help manipulating strings in R.

### Problem 1.

Using the 173 majors listed in fivethirtyeight.com’s College Majors dataset [https://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/], provide code that identifies the majors that contain either "DATA" or "STATISTICS"

#### Solution & Approach:
Upload dataset to GitHub repository, load it to R then proceed to the analysis. 

```{r}
# Load data from GitHub

college_majors<-read.csv("https://raw.githubusercontent.com/jnataky/RCharacter_manipulation/master/College_majors.csv")

# Check the columns names in dataframe
names(college_majors)
```

```{r}
# Count unique majors, to ensure no major is repeated

length(unique(college_majors[["Major"]]))
```


```{r}
# Identify the majors that contain either "DATA" or "STATISTICS" 
Major<-college_majors$Major
Major[grepl("DATA|STATISTICS", Major)]

```
#### Note: 

grepl() function is used to identify majors that contain either "DATA" or "STATISTICS".
What the function does is that it looks for ('DATA')(any character sequence)('STATISTICS') OR ('STATISTICS')(any character sequence)('DATA').

#### Conclusion: 

From the 173 majors, there are 3 majors containing either "DATA" or "STATISTICS".


### Problem 2.

Write code that transforms the data below:

[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"

[5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  

[9] "elderberry"   "lime"         "lychee"       "mulberry"    

[13] "olive"        "salal berry"

Into a format like this:

c("bell pepper", "bilberry", "blackberry", "blood orange", "blueberry", "cantaloupe", "chili pepper", "cloudberry", "elderberry", "lime", "lychee", "mulberry", "olive", "salal berry")


#### Solution & Approach:

Represent the given fruits as a string then convert it to a list following the pattern A-Z

```{r}
library(stringr)
# Write a string for fruits

fruits <-'[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"

[5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  

[9] "elderberry"   "lime"         "lychee"       "mulberry"    

[13] "olive"        "salal berry" '

# Convert str to a list with pattern p

p<-"[A-Za-z]+.?[A-Za-z]+"

fruits<- str_extract_all(fruits, p)

list_fruits <- str_c(fruits, sep = "", collapse = NULL)

list_fruits <-writeLines(list_fruits)

```


### Problem 3.

 
#### Approach:
Write in words then illustrate.

Describe, in words, what these expressions will match:


```{r}
# Define group of words, gr:
 gr <-c("eleven", "bob", "jujube", "church", "pepper", "coool", "monisnotnom")

```
 

* (.)\1\1, in str "(.)\\1\\1: It matches words with a character repeated 3 times.

```{r}
# Example

str_view(gr,"(.)\\1\\1", match = TRUE)

```


* "(.)(.)\\2\\1" : It matches words with pair of characters following by the reverse of that same pair of characters.
```{r}
# Example

str_view(gr,"(.)(.)\\2\\1", match = TRUE)
```


* (..)\1, in str "(..)\\1: It matches expressions with repeated pair of characters
```{r}
str_view(gr,"(..)\\1", match = TRUE)

```


* "(.).\\1.\\1": It matches words with 5 characters in which the first, third, and fifth character are the same.(The start, middle, and end character are the same)
```{r}
str_view(gr,"(.).\\1.\\1", match = TRUE)

```

* "(.)(.)(.).*\\3\\2\\1": It matches words in which the first 3 characters are followed by any characters then the reverse of the same first 3 characters.

```{r}

 
 str_view(gr, "(.)(.)(.).*\\3\\2\\1", match = TRUE)
 
```

### Problem 4.

 
#### Approach:
Use the same list as in problem 3 for illustration.


Construct regular expressions to match words that:

* Start and end with the same character.

```{r}

 
 str_view(gr, "^(.).*\\1$", match = TRUE)
 
```


* Contain a repeated pair of letters (e.g. "church" contains "ch" repeated twice.)

```{r}

 
 str_view(gr, "(..).*\\1", match = TRUE)
 
```

* Contain one letter repeated in at least three places (e.g. "eleven" contains three "e"s.)

```{r}

 
 str_view(gr, "(.).*\\1.*\\1", match = TRUE)
 
```