---
title: "DATA 607 ASSIGNMENT 5"
author: "Jered Ataky"
date: "`r Sys.Date()`"
output: 
  openintro::lab_report: default
  html_document:
    number_sections: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Librairies

```{r}
library(tidyverse)
library(kableExtra)
```

## Load data


```{r}

airlines <- read.csv("https://raw.githubusercontent.com/jnataky/DATA-607/master/A5_Data_transformation/airlines_dest.csv")
```


## Data Transformation

### Clean the data frame


```{r}

# Get rid of the empty row & reset the index

airlines <- na.omit(airlines)
row.names(airlines) <- NULL
```

```{r}
# Replace rows 2 and 4 by airlines names

airlines[2, "X"] <- "ALASKA"
airlines[4, "X"] <- "AM WEST"
```

```{r}
# Rename the columns names

airlines <- airlines %>%
  rename(carrier = X, status = X.1, los_angeles = Los.Angeles, phoenix = Phoenix, san_diego = San.Diego,
         san_francisco = San.Francisco, seattle = Seattle)

```


### Tidying the data frame


```{r}

# create new variable for destination and number of flights 

airlines <- airlines %>%
  gather("dest", "n_flights", 3:7)


```

```{r}

# Wider the data frame by adding variables "on time" and "delayed"

airlines <- airlines %>%
  pivot_wider(names_from = "status", values_from = "n_flights")

```

```{r}

# Table for tidy data


airlines %>%
  kbl(caption = "Airlines info", align = 'c') %>%
  kable_material(c("striped", "hover")) %>%
  row_spec(0, color = "indigo")
```

## Data Analysis

### Airlines performance
```{r}

# Create news variables

n_total <- airlines$delayed + airlines$`on time`
ontime_percent <- round((airlines$`on time` / n_total), 3)
delayed_percent <- round((airlines$delayed / n_total), 3)

airlines_df <- airlines %>%
  mutate(n_total, ontime_percent, delayed_percent)


airlines_df %>%
  kbl(caption = "Airlines more info", align = 'c') %>%
  kable_material(c("striped", "hover")) %>%
  row_spec(0, color = "indigo")
```


### Plotting the airlines performance

### Airlines on time comparison per city

```{r}

on_time1 <- airlines_df%>%
  group_by(dest, carrier) %>%
  summarise(ontime_percent)

on_time1 %>%
  kbl(caption = "On time performance per city", align = 'c') %>%
  kable_material(c("striped", "hover")) %>%
  row_spec(0, color = "indigo")


```

```{r}
# Plotting on time performance

ggplot(data = on_time1, aes(x = dest, y = ontime_percent, fill = carrier)) +
  geom_bar(stat = "identity", position = "dodge") +
  xlab("City") + ylab("On time % ") + ggtitle("Carriers on time performance per city")


```

### Note on airlines on-time performance per city

Alaska performs better than AM West in every single of the five cities.


### Overall on-time performance for both airlines


```{r}

on_time2 <- airlines_df%>%
  group_by(carrier) %>%
  summarise(ontime_wg_mean = mean(ontime_percent))

on_time2 %>%
  kbl(caption = "Overall on-time performance", align = 'c') %>%
  kable_material(c("striped", "hover")) %>%
  row_spec(0, color = "indigo")


```



```{r}

on_time2 <- airlines_df%>%
  group_by(carrier) %>%
  summarise(ontime_wg_mean = sum(`on time`) / sum(`on time`, delayed))

on_time2 %>%
  kbl(caption = "Overall on-time performance", align = 'c') %>%
  kable_material(c("striped", "hover")) %>%
  row_spec(0, color = "indigo")


```



```{r}
# Plotting overall on-time performance

ggplot(data = on_time2, aes(x = carrier, y = ontime_wg_mean)) +
  geom_bar(stat = "identity", width = 0.5, fill = "#b629a0") +
  xlab("Carrier") + ylab("On-time mean % ") + ggtitle("Overall on-time performance per carrier")
```


### Note on overall on-time performance per carrier

Overall AM West has more on-time flights than Alaska Air.


### A note on discrepancy


<style>
div.aquamarine { background-color:#7fffd4; border-radius: 10px; padding: 5px;}
</style>
<div class = "aquamarine">

When it comes to on-time performance per city for both airlines,
Alaska Airlines performs better but it is worse in overall on-time performance.
In per city performance, AM West doesn't perform better than Alaska might be
due to the number of flights its operates.



</div> \hfill\break


## Take Away

<style>
div.aquamarine { background-color:#7fffd4; border-radius: 10px; padding: 5px;}
</style>
<div class = "aquamarine">

The analyze has shown some discrepancy when comparing the performance between
per city and overall performance.

Before digging into the conclusion, let have a look on delays 
and analyze the overall number of flights per city, and see how it goes!



</div> \hfill\break




### Graphs and insights


```{r}

on_time3 <- airlines_df%>%
  group_by(dest, carrier) %>%
  summarise(ontime_percent)

# Plotting on time performance

ggplot(data = on_time3, aes(x = dest, y = delayed_percent, fill = carrier)) +
  geom_bar(stat = "identity", position = "dodge") +
  xlab("City") + ylab("Delayed  % ") + ggtitle("Carriers delay per city")


```


```{r}

on_time4 <- airlines_df%>%
  group_by(dest, carrier) %>%
  summarise(n_total)

# Plotting on time performance

ggplot(data = on_time4, aes(x = dest, y = n_total, fill = carrier)) +
  geom_bar(stat = "identity", position = "dodge") +
  xlab("City") + ylab("Number of flights ") + ggtitle("Number of flights per city")


```

### Conclusion


<style>
div.aquamarine { background-color:#7fffd4; border-radius: 10px; padding: 5px;}
</style>
<div class = "aquamarine">


Here's: Looking at the graphs above, we can see that Alaska operates more flights in Seattle and San Francisco.
In San Francisco, there is not much of difference in number of flights. This explain why Alaska Airlines beat AM West 
in per city performance. In the city where AM West operates more flights, it has significantly more flights than Alaska,
this could explain why there are more delays. But taking into consideration the number of flights AM West operates in these 5 cities, I will recommend AM West to review its reservation system. With that, it will perform way better than Alaska.
In the other hand, Alaska Airlines needs to review the internal problem that causes flights to delay.




</div> \hfill\break











