---
title: "DATA 607 ASSIGNMENT 7"
author: "Jered Ataky"
date: "`r Sys.Date()`"
output: 
  openintro::lab_report: default
  html_document:
    number_sections: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview


<style>
div.aquamarine { background-color:#7fffd4; border-radius: 10px; padding: 5px;}
</style>
<div class = "aquamarine">

I used Dreamwaever to create three different files in HTML, XML, and JSON
formats. These files are tables containing exact same information, which are 
the title, authors, edition, publisher, and publication year of three of 
my favorite books in learning data science. Note that two of my three books have 
two authors.

I will load these files into 3 three different data frames and then compare them
two by two to check their similarity and difference in R if there is any, and 
will share the findings.

</div> \hfill\break


## Libraries

```{r}

library(XML)
library(jsonlite)
library(RCurl)

library(tidyverse)
library(kableExtra)

```

## Data & Data frames

### Get URL

```{r}

html_url <- "https://raw.githubusercontent.com/jnataky/DATA-607/master/Web_technologies/books_html.html"

xml_url <- "https://raw.githubusercontent.com/jnataky/DATA-607/master/Web_technologies/books_xml.xml"

json_url <- "https://raw.githubusercontent.com/jnataky/DATA-607/master/Web_technologies/books_json.json"

```

### Load files

#### HTML

```{r}

file_html <- getURL(html_url)

df_html <- readHTMLTable(file_html, which =1)

# Kable for tidy table

df_html %>%
  kbl(caption = "My Favorite books") %>%
  kable_material(c("striped", "hover")) %>%
  row_spec(0, color = "indigo")

```


#### XML

```{r}

file_xml <- getURL(xml_url)

df_xml <- xmlToDataFrame(file_xml)


# Kable for tidy table

df_xml %>%
  kbl(caption = "My Favorite books") %>%
  kable_material(c("striped", "hover")) %>%
  row_spec(0, color = "indigo")

```



#### JSON

```{r}

df_json <- as.data.frame(fromJSON(json_url))

```

```{r}
# Change columns names to match others

names(df_json) <- c("Title", "Authors", "Edition", "Publisher", "Publication_Year")

# Kable for tidy table

df_json %>%
  kbl(caption = "My Favorite books") %>%
  kable_material(c("striped", "hover")) %>%
  row_spec(0, color = "indigo")

```

## Comparison


```{r}

# HTML & XML

all.equal(df_html, df_xml)

```

```{r}

# HTML & JSON

all.equal(df_html, df_json)

```


<style>
div.aquamarine { background-color:#7fffd4; border-radius: 10px; padding: 5px;}
</style>
<div class = "aquamarine">

Since the two first comparisons are true, 
we can say that all three frame have the same contents 
(after renaming to json data frame columns as the structure looked different
from the others). They look the same but the internal structure might be different.

Find more info on those formats in the Findings section below.

</div> \hfill\break


## Findings


<style>
div.aquamarine { background-color:#7fffd4; border-radius: 10px; padding: 5px;}
</style>
<div class = "aquamarine">

This practice on these different data exchange formats bring to this conclusion: 

The native structure of HTML does not naturally map into R objects. 
We can import HTML files as raw text, but this deprives us of the most useful features of these documents.
And XML is a more generic counterpart to HTML and a frequently used format to exchange data on the Web.
In the other hand, JSON is more lightweight due to its less verbose syntax and only allows a limited
set of data types that are compatible with many programming languages.

Source: Munzert, S. (2015). Automated Data Collection with R: A practical guide to web scraping and text mining. Chichester: Wiley.

</div> \hfill\break

- 